Manuscript Figures

### 2021-09-14 src/manuscript_figures.Rmd
remove pathway and top feature analysis from immunotherapy analysis figure (panels F and G)

supFig <- cowplot::plot_grid(cowplot::plot_grid(p1, p2$gtable, nrow = 1, 
                                                labels = c('A', 'B'), rel_widths = c(1.2, 2)), 
                          p4, nrow = 2, labels = c('', 'C'), rel_heights = c(1.2, 2))
# print p3 to supplements 
supFig_count <- supFig_count + 1
SuppFigures[[supFig_count]] <- supFig
fig.name <- paste0('SuppFigure', supFig_count, '.benchmarks.supp.pdf')
names(SuppFigures)[supFig_count] <- fig.name
ggsave(file.path(FigureFolder, fig.name), 
              plot = supFig, units = 'in', width = 8.27, height = 11.69)
SuppLegends[[supFig_count]] <- "A) Predictive power of latent factors in distinguishing cancer types comparing MAUI with PCA/MOFA/MCIA. Mean Balanced Accuracy computed on the test dataset (30% of held-out data) trained on 70% of the dataset using an Elastic Net multi-class model. B) Confusion matrix of predicted versus true labels of samples' cancer types. The color of each cell represents the percentage of samples from the cancer type in the rows that are classified into the cancer types on the column. Cells that deviate from the red color represent relatively lower
level of classification accuracy. C) Top maui LFs detected per cancer type. Size/color/opacity of the points represent the average neural activation value of the LF for the corresponding cancer type samples."
```

@@

```{r immunotherapy_response}
# immunotherapy response prediction analysis: 
# TODO: pass paths via args
settings_imtrp <- yaml::read_yaml('/data/local/buyar/arcas/pancancer_multiomics_manuscript/results/immunotherapy_response/Mariathasan_et_al/settings.yaml')
pathways <- readRDS('/data/local/buyar/arcas/subtyping_paper/data/GeneSets/msigdb.pathways.RDS')

@@

dt <- do.call(rbind, lapply(names(LFs_imtrp), function(tool) {
  M <- LFs_imtrp[[tool]]$mariathasan.gex
  # find LFs predictive of response to therapy 
  dt <- get_factor_specific_variables(M, colData$binaryResponse)[order(padj)]
  #dt <- compute_anova(M, colData$binaryResponse)
  dt$tool <- tool
  return(dt)
}))

dt[,.SD[which.min(padj)],by = c('tool', 'ref_cl')]
dt$tool <- toupper(gsub("_factors", "", dt$tool))
dt$ref_cl <- ifelse(dt$ref_cl == 'CR/PR', 'Responders', 'Non-responders')

@@

# top go terms for top features of top LF 
W <- FWs_imtrp$maui_factors$mariathasan.gex
feats <- data.table(W[,top_lf,drop=F], keep.rownames = T)[order(abs(get(top_lf)), decreasing = T)][1:100]
# top pathways for top lf
top_terms <- compute_geneset_enrichment(feats$rn, pathways)[order(pval)][padj < 0.05][1:5]
top_terms$term_wr <- stringr::str_wrap(gsub("_", " ", top_terms$term), width = 24)
p4 <- ggplot(top_terms, aes(x = reorder(term_wr, -pval), y = -log10(padj))) + 
  geom_bar(stat = 'identity') + 
  coord_flip() + labs(x = '') 
colnames(feats)[2] <- 'top_lf'
p5 <- ggplot(feats[1:10,], aes(x = reorder(rn, abs(top_lf)), y = abs(top_lf))) + 
  geom_bar(stat = 'identity') + coord_flip() + labs(x = '', y = 'Feature Importance (absolute)')

# correlation of top LF with tumor mutation burden
# get samples with both response and tmb annotation
dt <- do.call(rbind, lapply(names(LFs_imtrp), function(tool) {
  M <- LFs_imtrp[[tool]]$mariathasan.gex
  dt <- data.table(cor(M, colData$`Neoantigen burden per MB`, use = 'complete.obs'),
                   keep.rownames = T)
  dt$tool <- tool
  return(dt)
}))
dt <- dt[,.SD[which.max(abs(V1))], by = c('tool')]
dt$tool <- toupper(gsub("_factors", "", dt$tool))
p6 <- ggplot(dt, aes(x = tool, y = abs(V1))) + 
  geom_bar(stat = 'identity', aes(fill = tool), position = 'dodge') +
  coord_flip() + 
  labs(y = 'Correlation with Tumor Mutation Burden') + 
  theme(legend.title = element_blank(), axis.title.y = element_blank(), 
        plot.title = element_text(hjust = 0.5)) 
# immune phenotypes 
dt <- do.call(rbind, lapply(names(LFs_imtrp), function(tool) {
  M <- LFs_imtrp[[tool]]$mariathasan.gex
  # find LFs predictive of immune phenotypes 
  dt <- get_factor_specific_variables(M, colData$`Immune phenotype`)[order(padj)]
  dt$tool <- tool
  return(dt)
}))

dt[,.SD[which.min(padj)],by = c('ref_cl', 'tool')]
dt$tool <- toupper(gsub("_factors", "", dt$tool))
p7 <- ggplot(dt[,.SD[which.min(padj)],by = c('ref_cl', 'tool')], 
             aes(x = ref_cl, y = -log10(padj))) + 
  geom_bar(stat = 'identity', aes(fill = tool), position = 'dodge') +
  geom_hline(yintercept = -log10(0.05)) + 
  labs(x = 'Tumor Immune Phenotypes') + 
  coord_flip() + 
  theme(legend.title = element_blank(), axis.title.y = element_blank(), 
        plot.title = element_text(hjust = 0.5))

Fig <- cowplot::plot_grid(
  cowplot::plot_grid(p1, p6, p7, nrow = 1, labels = 'AUTO'),
  cowplot::plot_grid(p2$gtable, p3$plot, p4, p5, nrow = 2, labels = c('D', 'E', 'F', 'G')),
  nrow = 2, 
  rel_heights = c(1,2)
  ) 
fig_count <- fig_count + 1
Figures[[fig_count]] <- Fig
fig.name <- paste0('Figure', fig_count, '.immunotherapy_response.pdf')
names(Figures)[fig_count] <- fig.name
ggsave(filename = file.path(FigureFolder, fig.name), 
       plot = Fig, units = 'in', 
       width = 12, height = 14)
Legends[[fig_count]] <- "Predicting responders to anti-PD-L1 treatment in a cohort of Metastatic Urothelial
Cancers (N=348) using MAUI. A) Comparison of top latent factors predictive of complete/partial response (CR/PR) to the anti-PD-L1. B) Comparison of top latent factors with respect to their correlation with tumor mutation burden status of the samples in the cohort C) Comparison of top latent factors predictive of immune phenotypes (inflamed/excluded/desert). D) Top MAUI LFs predictive of response to anti-PD-L1 response groups. E) Survival stratification of the cohort based activation score of LF44 (high = LF44 > 1.4, low = LF44 < 1.4). F) Top pathways enriched among top 100 genes that contribute to the top latent factor (LF44). G) Top 10 genes contributing to LF44"
```

>>>>

supFig <- cowplot::plot_grid(cowplot::plot_grid(p1, p2$gtable, nrow = 1, 
                                                labels = c('A', 'B'), rel_widths = c(1.2, 2)), 
                          p4, nrow = 2, labels = c('', 'C'), rel_heights = c(1.2, 2))
# print p3 to supplements 
supFig_count <- supFig_count + 1
SuppFigures[[supFig_count]] <- supFig
fig.name <- paste0('SuppFigure', supFig_count, '.benchmarks.supp.pdf')
names(SuppFigures)[supFig_count] <- fig.name
ggsave(file.path(FigureFolder, fig.name), 
              plot = supFig, units = 'in', width = 8.27, height = 11.69)
SuppLegends[[supFig_count]] <- "A) Predictive power of latent factors in distinguishing cancer types comparing MAUI with PCA/MOFA/MCIA. Mean Balanced Accuracy computed on the test dataset (30% of held-out data) trained on 70% of the dataset using an Elastic Net multi-class model. B) Confusion matrix of predicted versus true labels of samples' cancer types. The color of each cell represents the percentage of samples from the cancer type in the rows that are classified into the cancer types on the column. Cells that deviate from the red color represent relatively lower
level of classification accuracy. C) Top maui LFs detected per cancer type. Size/color/opacity of the points represent the average neural activation value of the LF for the corresponding cancer type samples."

```

@@

```{r immunotherapy_response}
# immunotherapy response prediction analysis: 
# TODO: pass paths via args
settings_imtrp <- yaml::read_yaml('/data/local/buyar/arcas/pancancer_multiomics_manuscript/results/immunotherapy_response/Mariathasan_et_al/settings.yaml')

@@

dt <- do.call(rbind, lapply(names(LFs_imtrp), function(tool) {
  M <- LFs_imtrp[[tool]]$mariathasan.gex
  # find LFs predictive of response to therapy 
  dt <- get_factor_specific_variables(M, colData$binaryResponse)[order(padj)]
  dt$tool <- tool
  return(dt)
}))

dt$tool <- toupper(gsub("_factors", "", dt$tool))
dt$ref_cl <- ifelse(dt$ref_cl == 'CR/PR', 'Responders', 'Non-responders')

@@

# correlation of top LF with tumor mutation burden
# get samples with both response and tmb annotation
dt <- do.call(rbind, lapply(names(LFs_imtrp), function(tool) {
  M <- LFs_imtrp[[tool]]$mariathasan.gex
  dt <- data.table(cor(M, colData$`Neoantigen burden per MB`, use = 'complete.obs'),
                   keep.rownames = T)
  dt$tool <- tool
  return(dt)
}))
dt <- dt[,.SD[which.max(abs(V1))], by = c('tool')]
dt$tool <- toupper(gsub("_factors", "", dt$tool))
p4 <- ggplot(dt, aes(x = tool, y = abs(V1))) + 
  geom_bar(stat = 'identity', aes(fill = tool), position = 'dodge') +
  coord_flip() + 
  labs(y = 'Correlation with Tumor Mutation Burden') + 
  theme(legend.title = element_blank(), axis.title.y = element_blank(), 
        plot.title = element_text(hjust = 0.5)) 
# immune phenotypes 
dt <- do.call(rbind, lapply(names(LFs_imtrp), function(tool) {
  M <- LFs_imtrp[[tool]]$mariathasan.gex
  # find LFs predictive of immune phenotypes 
  dt <- get_factor_specific_variables(M, colData$`Immune phenotype`)[order(padj)]
  dt$tool <- tool
  return(dt)
}))

dt[,.SD[which.min(padj)],by = c('ref_cl', 'tool')]
dt$tool <- toupper(gsub("_factors", "", dt$tool))
p5 <- ggplot(dt[,.SD[which.min(padj)],by = c('ref_cl', 'tool')], 
             aes(x = ref_cl, y = -log10(padj))) + 
  geom_bar(stat = 'identity', aes(fill = tool), position = 'dodge') +
  geom_hline(yintercept = -log10(0.05)) + 
  labs(x = 'Tumor Immune Phenotypes') + 
  coord_flip() + 
  theme(legend.title = element_blank(), axis.title.y = element_blank(), 
        plot.title = element_text(hjust = 0.5))

Fig <- cowplot::plot_grid(
  cowplot::plot_grid(p1, p4, p5, nrow = 1, labels = 'AUTO'),
  cowplot::plot_grid(p2$gtable, p3$plot, nrow = 1, labels = c('D', 'E')),
  nrow = 2, 
  rel_heights = c(1,2)
  ) 
fig_count <- fig_count + 1
Figures[[fig_count]] <- Fig
fig.name <- paste0('Figure', fig_count, '.immunotherapy_response.pdf')
names(Figures)[fig_count] <- fig.name
ggsave(filename = file.path(FigureFolder, fig.name), 
       plot = Fig, units = 'in', 
       width = 12, height = 14)
Legends[[fig_count]] <- "Predicting responders to anti-PD-L1 treatment in a cohort of Metastatic Urothelial
Cancers (N=348) using MAUI. A) Comparison of top latent factors predictive of complete/partial response (CR/PR) to the anti-PD-L1. B) Comparison of top latent factors with respect to their correlation with tumor mutation burden status of the samples in the cohort C) Comparison of top latent factors predictive of immune phenotypes (inflamed/excluded/desert). D) Top MAUI LFs predictive of response to anti-PD-L1 response groups. E) Survival stratification of the cohort based activation score of LF44 (high = LF44 > 1.4, low = LF44 < 1.4)."
```
